---
title: "assignment 1"
author: "Nico"
date: "6/22/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dataset Information
This dataset has information regarding the different college majors and the respective statistics including basic earnings and labor force information. The data was collected from the American Community Survey 2010-2012 Public Use Microdata series. Based on the information here, I think it would be a good clustering use case. We can see how similar some of these majors are based on the various features that are provided in the dataset. We might be surprised to see the similarities between some of these majors that aren't so common. Its currently in a csv file so it will be pretty easy to load into R since its comma delimited. 

```{r}
library(rmarkdown)
library(tidyr)
library(dplyr)
```

# Import file into Rstudio
We first download the the raw csv file and call readcsv to read the data into our environment where we assign it to recentgrads variable.

```{r}
download.file(url = "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/recent-grads.csv", destfile = "recentgrads.csv")
recentgrads<-read.csv("recentgrads.csv", stringsAsFactors = FALSE)
```

Just to get a preview of our data we will use the head command to view the first 5 rows of our dataframe.
```{r}
head(recentgrads,5)
```

# Date Exploration
view the column names within our dataframe
```{r}
colnames(recentgrads)
```

NExt let's filter our dataset to the only the columns we need using the select command from dplyr library. I also renamed our dataframe to df to make it easier.
```{r}
df <- recentgrads %>% 
  select(Major,Major_category,Total,Unemployed, Full_time)
head(df)
```
We can also rename some of these columns to give a better description
```{r}
colnames(df)[colnames(df) == 'Total'] <- 'Total_people'
```
# Data characteristics
This altered genderdata dataset has `r ncol(df)` columns and `r nrow(df)` rows. We can we can see the types of data in a more concise format through the code below. : 
```{r}
summary.default(df)
```
Next, we will build a table with the column details.
```{r}
text_tbl <- data.frame(
  Names = c("Major", "Major_category", "Total_people", "Unemployed", "Full_time"),
  Description = c(
    "Major description",
    "Category of major from Carnevale et al", 
    "Total number of people with major",
    "Number unemployed (ESR == 3)",
    "Employed 35 hours or more")
)

text_tbl

```

# Summary statistics
Using the summary function in R we are able to get the summaries of the each column in our dataset. The summary function also shows any null values for each column where for our case Total_people column has 1 null value. Summary function by default will exclude the null value when calculating for the minimum, maximum and mean. 
```{r}
Summary_details<-summary(df)
colSums(is.na(df))
Summary_details
```



